Overview and Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE)

### 2014-02-20 WSSSPE1_report/WSSSPE1_summary_report.tex
small update

These categories have some overlap, and in some \note{most?} cases, the
case studies don't cleanly fit into a single category.
Our classification is based on two criteria: first, the character
of a software project that dominates and best fits, and, second,
the theme of the workshop discussion in which the paper discussing
the project was placed.

@@

\subsection{Utility Software}
Utility software potentially has the broadest community impact of the three
categories. This is
especially relevant in the modern era of social computing. In the context of
this paper, a utility software project could be defined as general purpose software
applicable to one or many generic tasks and/or enabling other software to run
by providing a suitable environment. Examples are collaborative development
framework such as \emph{GitHub} and \emph{Bitbucket}, workflow and scripting
frameworks such as \emph{Galaxy}, \emph{Swift}, and \emph{VisTrails},
and visualization frameworks such as \emph{visIT}. Often projects start with
specific, small set of requirements and users in mind but expand to provide
more generic functionality and appeal to a wider user community.

>>>>

These categories have some overlap, and in some \note{most?} cases, the
case studies do not cleanly fit into a single category.
Our classification is based on two criteria: first, the character
of a software project that dominates and best fits, and, second,
the theme of the workshop discussion in which the paper discussing
the project was placed.

@@

\subsection{Utility Software}
Utility software potentially has the broadest community impact of the three
categories. This is especially relevant in the modern era of social computing. In the context of
this paper, a utility software project could be defined as general purpose software
applicable to one or many generic tasks and/or enabling other software to run
by providing a suitable environment. Examples are collaborative development
framework such as \emph{GitHub} and \emph{Bitbucket}, workflow and scripting
frameworks such as \emph{Galaxy}, \emph{Swift}, and \emph{VisTrails},
and visualization frameworks such as \emph{visIT}. Often projects start with
specific, small set of requirements and users in mind but expand to provide
more generic functionality and appeal to a wider user community. Owing to its generic nature, utility software finds a wider users communities which in turn leads to a higher participation in development. Consequently, the process becomes user-driven and self sustaining. Software developed in such scenario often benefits from community driven development, deployment and maintenance best practices.

For instance the \emph{Galaxy} project has significantly benefitted from user community participation in its development. It follows agile software development practices and implements standard practices such as test-driven development and socialized bug managing practices via \emph{trello}. Galaxy \emph{histories} and \emph{toolshed} offers easy community sharing of data and tools further promoting a collaborative environment.
